% ********** Abstract **********
\chapter*{Abstract}

Powering a device using solar generated energy can be difficult, especially when that device is
meant to function constantly over a long period of time. In this thesis we will present an
architecture for energy harvesting wireless sensor networks that can be used to develop solar
powered applications. It will cover both hardware as well as the software needs in an energy
harvesting wireless sensor network. The hardware will be composed of a new low power node designed to be a powerful development
platform and an efficient energy harvesting module. The software is designed to efficiently use the
stored energy by implementing a lightweight algorithm for scheduling data transmission.




\textbf{Keywords} Wireless Sensor Networks, Sparrow, Sparrow R, energy harvesting, scheduling,
solar powered, dynamic scheduling

\chapter*{Acknowledgements}


Foremost, I would like to express my gratitude to my advisor Dan Tudose for the continuous support that he provided.

Besides my advisor, I would also like to give special thanks to Dan Dragomir for providing corrections to this work.

My sincere thanks also goes to Adriana Drăghici, Alexandru Corneliu Olteanu and Andrei Voinescu for their insightful comments  and moral support.

I thank my fellow labmates from the Robolab robotics group: Tudor Vișan, Andrei Mușat and Andrei Vasiliu for the sleepless nights we were working together before deadlines, and for all the fun we had in the past four years.

Last but not the least, I would like to thank my girlfriend and my family who have shown understanding and have given me the moral support needed.

% ********** End of Abstract **********
